Nitric oxide feedback to ciliary photoreceptor cells gates a UV avoidance circuit

### 2023-06-09 Manuscript/Manuscript.Rmd
Ms: methods, renamed suppl file
Update methids

We used a single-cell RNA sequencing dataset from *Platynereis* larvae [@Achim2015] that was further filtered to 107 cells [@williams2017].
We normalized raw read-count data to TPM and converted to log10.
To calculate % expression, we used the sum of the expression levels in the 107 cells for each gene.
The total TPM of each gene between the samples was used to calculate the percentage of expressed genes.
We processed the data in Python and plotted in R.
Individual cells were identified by suits of well-characterised marker genes and spatial mapping [@Achim2015].
The total TPM values for each gene and the percentage of expressed genes are available in **Supplementary file 1**.

@@

To measure cGMP production in a cell culture assay we used the Green cGull reporter [@Matsuda2017].
Full-length *NIT-GC1* and *NIT-GC2* coding sequences were amplified by PCR from a *Platynereis dumerilii* cDNA library and cloned into the pcDNA3.1(+) vector with the T2A self-cleaving sequence.
For the assay, COS-7 cells (Angio-proteomie, CAT no. cAP-0203) with a low endogenous level of soluble guanylate cyclase activity were used.
Cells were maintained at 37 °C in 35 mm dishes (Nunc™ Glass Bottom Dishes) containing 3 ml of DMEM, high glucose glutamax medium (Thermo; Cat. No. 10566016) supplemented with 10% fetal bovine serum (Thermo; Cat. No. 10082147).
Upon reaching approximately 85% confluency, cells were transfected with the pcDNA3.1(+)-Green-cGull-T2A-NITGC1 plasmid.
Transfections were carried out with 150 ng of each plasmid and 0.3 μl of Lipofectamine 3000 Reagent (invitrogen; Cat. No. L3000001).
Two days post-transfection, the culture medium was substituted for fresh DMEM-medium. 
For single-wavelength imaging experiments, cells in 35-mm dishes were washed twice and imaged in modified Ringer's buffer (140 mM NaCl, 3.5 mM KCl, 0.5 mM NaH~2~PO~4~ , 0.5 mM S-3 MgSO~4~, 1.5 mM CaCl~2~ , 10 mM HEPES, 2 mM NaHCO~3~ and 5 mM glucose).
Dishes were mounted on a stage heated to 37 °C and imaging was done on an inverted confocal microscope (LSM880, Carl Zeiss GmbH, Jena) equipped with an oil-immersion objective lens (UApo/340, 40x, NA = 0.17).
Images were acquired using a xenon lamp, a 460-495 nm excitation filter, a 505-nm dichroic mirror and a 510-550-nm emission filter (Carl Zeiss GmbH, Jena).
The exposure time of the EM-CCD camera was controlled by the ZEN software (Carl Zeiss GmbH, Jena).
Images were acquired every 15 s for 10 min and stimulation was initiated 2 min after starting image acquisition.
Imaging data analysis was done with ImageJ (National Institutes of Health, Bethesda, MD, USA).
As NO donor, we used S-Nitroso-N-acetyl-D, L-penicillamine (SNAP, Sigma-Aldrich, St. Louis, MO, USA).

>>>>

We used a single-cell RNA sequencing dataset from *Platynereis* larvae [@Achim2015] that was further filtered to 107 cells [@williams2017].
We normalized raw read-count data to TPM and converted to log10.
To calculate % expression, we used the sum of the expression levels in the 107 cells for each gene.
The total TPM of each gene between the samples was used to calculate the percentage of expressed genes.
We processed the data in Python and plotted in R.
Individual cells were identified by suits of well-characterised marker genes and spatial mapping [@Achim2015].
The total TPM values for each gene and the percentage of expressed genes are available in **Supplementary file 7**.

@@

To measure cGMP production in a cell culture assay we used the Green cGull reporter [@Matsuda2017].
Full-length *NIT-GC1* and *NIT-GC2* coding sequences were amplified by PCR from a *Platynereis dumerilii* cDNA library and cloned into the pcDNA3.1(+) vector with the T2A self-cleaving sequence.
For the assay, COS-7 cells (Angio-proteomie, CAT no. cAP-0203) with a low endogenous level of soluble guanylate cyclase activity were used.
Cells were maintained at 37 °C in 35 mm dishes (Nunc™ Glass Bottom Dishes) containing 3 ml of DMEM, high glucose glutamax medium (Thermo; Cat. No. 10566016) supplemented with 10% fetal bovine serum (Thermo; Cat. No. 10082147).
Upon reaching approximately 85% confluency, cells were transfected with the pcDNA3.1(+)-Green-cGull-T2A-NITGC1 plasmid.
Transfections were carried out with 150 ng of each plasmid and 0.3 μl of Lipofectamine 3000 Reagent (invitrogen; Cat. No. L3000001).
Two days post-transfection, the culture medium was substituted for fresh DMEM-medium. 
For single-wavelength imaging experiments, cells in 35-mm dishes were washed twice and imaged in modified Ringer's buffer (140 mM NaCl, 3.5 mM KCl, 0.5 mM NaH~2~PO~4~ , 0.5 mM S-3 MgSO~4~, 1.5 mM CaCl~2~ , 10 mM HEPES, 2 mM NaHCO~3~ and 5 mM glucose).
Dishes were mounted on a stage heated to 37 °C and imaging was done on an inverted confocal microscope (LSM880, Carl Zeiss GmbH, Jena) equipped with an oil-immersion objective lens (UApo/340, 40x, NA = 0.17).
The exposure time of the EM-CCD camera was controlled by the ZEN software (Carl Zeiss GmbH, Jena).
Images were acquired every 15 s for 10 min and stimulation was initiated 2 min after starting image acquisition.
Imaging data analysis was done with ImageJ (National Institutes of Health, Bethesda, MD, USA).
As NO donor, we used S-Nitroso-N-acetyl-D, L-penicillamine (SNAP, Sigma-Aldrich, St. Louis, MO, USA).
